---
title: "PROJECT 2 RCODE"
author: "Miguel Melgarejo"
date: "3/17/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/mmelgare/Desktop/R Thangs/560 Data Mining/Datasets"))
```

Install packages:
```{r eval=FALSE}
install.packages("mlbench")
```

Load packages: 
```{r}
library(e1071)
library(klaR)
library(nnet)
library(MASS)
library(rpart)
library(MASS)
library(klaR)
library(randomForest)
library(mlbench)
library(caret)
```

Load data: 
```{r}
data(BreastCancer)

BreastCancer <- na.omit(BreastCancer)
BreastCancer <- BreastCancer[-1] #removing the unique identifiers to not confuse the models

mydata <- cbind(BreastCancer[10],BreastCancer[1:9]) #moving Class to the front so that it is easier to visualize 

```

Data type transformation: 
```{r}
#Convert the variables from factors to integers

mydata$Class<-as.integer(BreastCancer$Class)
mydata$Cl.thickness<-as.integer(BreastCancer$Cl.thickness)
mydata$Cell.size<-as.integer(BreastCancer$Cell.size)
mydata$Cell.shape<-as.integer(BreastCancer$Cell.shape)
mydata$Marg.adhesion<-as.integer(BreastCancer$Marg.adhesion)
mydata$Epith.c.size<-as.integer(BreastCancer$Epith.c.size)
mydata$Bare.nuclei<-as.integer(BreastCancer$Bare.nuclei)
mydata$Bl.cromatin<-as.integer(BreastCancer$Bl.cromatin)
mydata$Normal.nucleoli<-as.integer(BreastCancer$Normal.nucleoli)
mydata$Mitoses<-as.integer(BreastCancer$Mitoses)


sum(is.na(BreastCancer)) #check to make sure na's have been removed 
str(mydata$Class) #check structure to verify data types have been updated
```


Classifier 1: SVM
```{r}
mysvm <- svm(Class ~ ., BreastCancer)
mysvm.pred <- predict(mysvm, BreastCancer)
table(mysvm.pred,BreastCancer$Class)
```

Classifier 2: Naive Bayes
```{r}
mynb <- NaiveBayes(Class ~ ., BreastCancer)
mynb.pred <- predict(mynb,BreastCancer)
table(mynb.pred$class,BreastCancer$Class)
```

Classifier 3: Neural Net 
```{r}
mynnet <- nnet(Class ~ ., BreastCancer, size=1)
mynnet.pred <- predict(mynnet,BreastCancer,type="class")
table(mynnet.pred,BreastCancer$Class)
```

Classifier 4: Decision Tree
```{r}
mytree <- rpart(Class ~ ., BreastCancer)
plot(mytree); text(mytree) # in "Breastcancer_tree.ps"
summary(mytree)
mytree.pred <- predict(mytree,BreastCancer,type="class")
table(mytree.pred,BreastCancer$Class)
```

Classifier 5: Leave-1-Out Cross Validation (LOOCV)
```{r}
ans <- numeric(length(BreastCancer[,1]))
for (i in 1:length(BreastCancer[,1])) {
  mytree <- rpart(Class ~ ., BreastCancer[-i,])
  mytree.pred <- predict(mytree,BreastCancer[i,],type="class")
  ans[i] <- mytree.pred
}
ans <- factor(ans,labels=levels(BreastCancer$Class))
table(ans,BreastCancer$Class)
# The same as above in this case
```

Classifier 6: Quadratic Discriminant Analysis
```{r}
myqda <- qda(Class ~ ., mydata)
myqda.pred <- predict(myqda, mydata)

table(myqda.pred$class,mydata$Class)
```

Classifier 7: Regularized Discriminant Analysis
```{r}
myrda <- rda(Class ~ ., BreastCancer)
myrda.pred <- predict(myrda, BreastCancer)
table(myrda.pred$class,BreastCancer$Class)
```

Classifier 8: Random Forests
```{r}
myrf <- randomForest(Class ~ .,BreastCancer)
myrf.pred <- predict(myrf, BreastCancer)
table(myrf.pred, BreastCancer$Class)
```

Combining models: 
```{r}
#combining all the models into a new dataframe so that ensemble can take place. 

combine.classes <- data.frame(myrf.pred, myrda.pred$class, #myqda.pred, 
                              mytree.pred, mynnet.pred, mysvm.pred, mynb.pred$class)

head(combine.classes) #view the dataframe to verify the structure and values the new datafram contains
head(myrf.pred)
head(myrda.pred)

```

Making models binary:
```{r}
#Convert the values to binary (o, 1) so that the models can be summed across

combine.classes$myrf.pred<-ifelse(combine.classes$myrf.pred=="benign", 0, 1)
combine.classes[,2]<-ifelse(combine.classes[,2]=="benign", 0, 1)
combine.classes[,3]<-ifelse(combine.classes[,3]=="benign", 0, 1)
combine.classes[,4]<-ifelse(combine.classes[,4]=="benign", 0, 1)
combine.classes[,5]<-ifelse(combine.classes[,5]=="benign", 0, 1)
combine.classes[,6]<-ifelse(combine.classes[,6]=="benign", 0, 1)
```

Majority vote: 
```{r}
#conduct the rowsums function and and save to a new dataframe

majority.vote=rowSums(combine.classes)
head(majority.vote)
```

Produce final table: 
```{r}


combine.classes[,7]<-rowSums(combine.classes)
combine.classes[,8]<-ifelse(combine.classes[,7]>=4, "malignant", "benign")
table(combine.classes[,8], BreastCancer$Class)

```

## R Markdown